Identifying the L1 of non-native writers: the CMU-Haifa system\\[1em] {\large Chris Dyer$^\ast$ Manaal Faruqui$^\ast$ Noam Ordan$^\dagger$ Nathan Schneider$^\ast$\\ Yulia Tsvetkov$^\ast$ Naama Twitto$^\dagger$ Shuly Wintner$^\dagger$}\\[-3em]

### 2013-04-07 paper/paper.tex
merge analysis

\documentclass[11pt,letterpaper]{article}
\usepackage[linkcolor=blue]{hyperref}	% must precede ACL style
\usepackage{relsize} % relative font sizes (e.g. \smaller). must precede ACL style
\usepackage{naaclhlt2013}
\usepackage{times}
\usepackage{latexsym}

@@

\newcommand{\feat}[1]{\textsmaller[.5]{\textsf{#1}}} % code for a feature group

@@

\section{Introduction}
\label{sec:intro}
The task we address in this work is identifying the native language
(\emph{L1}) of non-native English authors. More specifically, given a
dataset of short English essays
\citep{blanchard-tetreault-higgins-cahill-chodorow:2013:TOEFL11-RR},
composed as part of the \emph{Test of English as a Foreign Language
  (TOEFL)} by authors whose native language is one out of~11 possible
languages (Arabic, Chinese, French, German, Hindi, Italian, Japanese,
Korean, Spanish, Telugu, and Turkish), our task is to identify that
language.

@@

\begin{compactdesc}
\item[POS $n$-grams] Part-of-speech $n$-grams were used in various
  text-classification tasks.
\item[Prompt] Since the prompt contributes information on the domain,
  it is likely that some words (and, hence, character sequences) will
  occur more frequently with some prompts than with others. We
  therefore use the prompt ID in conjunction with other features.
\item[Document length] The number of tokens in the text is highly
  correlated with the author's level of fluency, which in turn is
  correlated with the author's L1.
\item[Pronouns] The use of pronoun varies greatly among different
  authors. We use the same list of~25 English pronouns that
  \citet{vered:noam:shuly} use for identifying translationese.
\item[Punctuation] Similarly, different languages use punctuation
  differently, and we expect this to taint the use of punctuation in
  non-native texts.
\item[Passives] English uses passive voice more frequently than other
  languages. Again, the use of passives in L2 can be correlated with
  the author's L1.
\item[Positional token frequency] The choice of the first and last few
  words in a sentence is highly constrained, and may be significantly
  influenced by the author's L1.
\item[Cohesive markers] These are 40 function words (and short
  phrases) that have a strong discourse function in texts (`however',
  `because', `in fact', etc.) Translators tend to spell out implicit
  utterances and render them explicitly in the target text
  \citep{Blum-Kulka:1986}. We use the list
  of~\citet{vered:noam:shuly}.
\item[Cohesive verbs] This is a list of manually compiled verbs that
  are used, like cohesive markers, to spell out implicit utterances
  (`indicate', `imply', `contain', etc.)
\item[Function words] Frequent tokens, which are mostly function
  words, have been used successfully for various text classification
  tasks. \citet{koppel-ordan:2011:ACL-HLT2011} define a list of 400
  such words, of which we only use~100 (using the entire list was not
  significantly different).
\item[Contextual function words] To further capitalize on the ability
  of function words to discriminate, we define pairs consisting of a
  function word from the list mentioned above, along with the POS tag
  of its adjacent word. This feature captures patterns such as verbs
  and the preposition or particle immediately to their right, or nouns
  and the determiner that precedes them. We also define 3-grams
  consisting of one or two function words and the POS tag of the third
  word in the 3-gram.
\item[Lemmas] The content of the text is not considered a good
  indication of the author's L1, but many text categorization tasks
  use lemmas (more precisely, the stems produced by the tagger) as
  features approximating the content.
\item[Misspelling features] Clearly, the spelling errors that learners
  make in English depend on the phonological properties of their
  L1. \sw{???}
\item[Restored tags] We focus on three important token classes defined
  above: punctuation marks, function words and cohesive verbs. We
  first remove words in these classes from the texts, and then recover
  the most likely hidden tokens \yt{To Shuly: I removed 'POS'. We recover words, not POS tags, I used the word 'tags' to distinguish between words to recover and all other words. Probably the word 'tag' is misleading, I changed it to 'token'} in a sequence of words, according to
  an $n$-gram language model trained on all essays in the training
  corpus corrected with a spell checker and containing both words and
  hidden tokens. This feature should capture specific words or
  punctuation marks that are consistently omitted (deletions), or
  misused (insertions, substitutions). To restore hidden tokens we use
  the \texttt{hidden-ngram} utility provided in SRILM
  \citep{stolcke02srilm}.
\item[Brown clusters] \sw{???}
\end{compactdesc}

@@

\section{Results}
\label{sec:results}
The full model that we used to classify the test set combines all
features listed in \Tref{tbl:addfeats}. Using all these features, the
accuracy on the development set is \sw{???}, and on the test set it
is~81.5\%. \Tref{tbl:matrix} lists the confusion matrix on the test
set, as well as precision, recall and $F_1$-score for each L1.

@@

\sw{Analysis? Error analysis? Observations?}

>>>>

\documentclass[11pt,letterpaper]{article}
\usepackage[linkcolor=blue]{hyperref}	% must precede ACL style
\usepackage{relsize} % relative font sizes (e.g. \smaller). must precede ACL style
\usepackage{naaclhlt2013}
\usepackage{times}
\usepackage{latexsym}
\usepackage{tipa}

@@

\newcommand{\feat}[1]{\textsmaller[.5]{\textsf{#1}}} % code for a feature group
\newcommand{\textnl}{\textsl}

@@

\section{Introduction}
\label{sec:intro}
The task we address in this work is identifying the native language
(\emph{L1}) of non-native English (\emph{L2}) authors. More specifically, given a
dataset of short English essays
\citep{blanchard-tetreault-higgins-cahill-chodorow:2013:TOEFL11-RR},
composed as part of the \emph{Test of English as a Foreign Language
  (TOEFL)} by authors whose native language is one out of~11 possible
languages (Arabic, Chinese, French, German, Hindi, Italian, Japanese,
Korean, Spanish, Telugu, and Turkish), our task is to identify that
language.

@@

\begin{compactdesc}
\item[POS $n$-grams] Part-of-speech $n$-grams were used in various
  text-classification tasks.
\item[Prompt] Since the prompt contributes information on the domain,
  it is likely that some words (and, hence, character sequences) will
  occur more frequently with some prompts than with others. We
  therefore use the prompt ID in conjunction with other features.
\item[Document length] The number of tokens in the text is highly
  correlated with the author's level of fluency, which in turn is
  correlated with the author's L1.
\item[Pronouns] The use of pronouns varies greatly among different
  authors. We use the same list of~25 English pronouns that
  \citet{vered:noam:shuly} use for identifying translationese.
\item[Punctuation] Similarly, different languages use punctuation
  differently, and we expect this to taint the use of punctuation in
  non-native texts. Of course, character $n$-grams subsume this feature.
\item[Passives] English uses passive voice more frequently than other
  languages. Again, the use of passives in L2 can be correlated with
  the author's L1.
\item[Positional token frequency] The choice of the first and last few
  words in a sentence is highly constrained, and may be significantly
  influenced by the author's L1.
\item[Cohesive markers] These are 40 function words (and short
  phrases) that have a strong discourse function in texts (\textnl{however},
  \textnl{because}, \textnl{in fact}, etc.) Translators tend to spell out implicit
  utterances and render them explicitly in the target text
  \citep{Blum-Kulka:1986}. We use the list
  of~\citet{vered:noam:shuly}.
\item[Cohesive verbs] This is a list of manually compiled verbs that
  are used, like cohesive markers, to spell out implicit utterances
  (\textnl{indicate}, \textnl{imply}, \textnl{contain}, etc.)
\item[Function words] Frequent tokens, which are mostly function
  words, have been used successfully for various text classification
  tasks. \citet{koppel-ordan:2011:ACL-HLT2011} define a list of 400
  such words, of which we only use~100 (using the entire list was not
  significantly different). Note that pronouns are included in this list.
\item[Contextual function words] To further capitalize on the ability
  of function words to discriminate, we define pairs consisting of a
  function word from the list mentioned above, along with the POS tag
  of its adjacent word. This feature captures patterns such as verbs
  and the preposition or particle immediately to their right, or nouns
  and the determiner that precedes them. We also define 3-grams
  consisting of one or two function words and the POS tag of the third
  word in the 3-gram.
\item[Lemmas] The content of the text is not considered a good
  indication of the author's L1, but many text categorization tasks
  use lemmas (more precisely, the stems produced by the tagger) as
  features approximating the content.
\item[Misspelling features] Clearly, the spelling errors that learners
  make in English depend on the phonological properties of their
  L1. \sw{???}
\item[Restored tags] We focus on three important token classes defined
  above: punctuation marks, function words and cohesive verbs. We
  first remove words in these classes from the texts, and then recover
  the most likely hidden tokens in a sequence of words, according to
  an $n$-gram language model trained on all essays in the training
  corpus corrected with a spell checker and containing both words and
  hidden tokens. This feature should capture specific words or
  punctuation marks that are consistently omitted (deletions), or
  misused (insertions, substitutions). To restore hidden tokens we use
  the \texttt{hidden-ngram} utility provided in SRILM
  \citep{stolcke02srilm}.
\item[Brown clusters] \sw{???}
\end{compactdesc}

@@

\section{Results}
\label{sec:results}
The full model that we used to classify the test set combines all
features listed in \Tref{tbl:addfeats}. Using all these features, the
accuracy on the development set is \sw{???}, and on the test set it
is~81.5\%. \Tref{tbl:matrix} lists the confusion matrix on the test
set, as well as precision, recall and $F_1$-score for each L1.
The largest class of errors was predicting Telugu where the correct label 
was Hindi\nss{or vice versa?}---this happened 18 times.

@@

Production of L2 texts, not unlike translating from L1 to L2, involves
a tension between the imposing models of L1 (and the source text), on
the one hand, and a set of cognitive constraints resulting from the
efforts to generate the target text, on the other. The former is
called \emph{interference} in Translation Studies \citep{Toury:1995}
and \emph{interlanguage} in second language acquisition
\citep{Selinker1972}.

\citet{vered:noam:shuly} designed 32 classifiers to test the validity
of the forces acting on translated texts, and found that interference
consistently yielded the best performing classifiers. And indeed, in
this work too, which replicates some of their classifiers, we find
again that fingerprints of the source language are dominant in the
makeup of L2 texts.

The main difference, however, between texts translated by
professionals and the texts we address here, is that more often than
not professional translators translate into their mother tongue,
whereas L2 writers write out of their mother tongue by definition. So
interference is ever more exaggerated in their case, for example, also
phonologically \citep{tsur-rappoport:2007:Cognitive-2007}.

We illustrate this with some examples from Arabic native speakers. The
character sequence \textnl{alot} is overrepresented in Arabic L2
texts. Arabic has no indefinite article and we speculate that Arabic
speakers conceive \textnl{a lot} as a single word; the Arabic
equivalent for \textnl{a lot} is used adverbially like an \textnl{-ly}
suffix in English. For the same reason, another prominent feature is a
missing definite article before nouns and adjectives . Additionally,
Arabic, being an Abjad language, rarely indicates vowels, and indeed
we find many missing \textnl{e}-s and \textnl{i}-s in the texts of
Arabic speakers. Phonologically, because Arabic conflates
/\textipa{I}/ and /\textipa{@}/ into /i/ (at least in Modern Standard
Arabic), we see that many \textnl{e}-s are indeed substituted for
\textnl{i}-s in these texts.

German overuses hyphens in two interesting ways. German can
notoriously use relative clauses freely, and such constructions
frequently occur between hyphens in the dataset, as in \textnl{any
  given rational being -- let us say Immanual Kant -- we find
  that}. Another overuse of hyphens stems from compounding, another
facet of German, for example in \textnl{well-known},
\textnl{community-help}, \textnl{spare-time}, \textnl{football-club},
etc. Many of these reflect an effort to both connect and separate
connected forms in the original (e.g., \textnl{Fussballklub}, which in
English would be more naturally rendered as \textnl{football club}).
Another unexpected feature of German is a frequent substitution of the
letter \textnl{y} for \textnl{z} (and vice versa), most probably
triggered by their switched positions on German keyboards.

The word \textnl{that} occurs more frequently in the texts of German
L1 speakers, perhaps because in English it is optional in relative
clauses whereas in German it is not.  Last, \textnl{often} is
overused; being a cognate of the German \textnl{oft} it is not
cognitively expensive to retrieve it. Spanish, on the other hand,
literally translates \textnl{muchas veces} into \textnl{many times},
which is similarly overused on the dataset. \sw{any Spanish speaker
  who could validate this expression?}

Other informative features include substitutions of \textnl{r}-s and
\textnl{l}-s in the texts of Japanese authors, for obvious reasons;
and the characters \textnl{r} and \textnl{s} are important in Chinese
and Spanish, respectively, for reasons that are unclear to
us. Similarly, the word \textnl{then} is dominant in the texts of
Hindi speakers. Finally, it is clear that authors refer to their
native cultures (and, consequently, native languages and countries);
the strings \textnl{Turkish}, \textnl{Korea}, and \textnl{Ita} were
dominant in the texts of Turkish, Korean and Italian native speakers,
respectively.
